---
title: "Penguin Flipper Lengths ANOVA"
author: "Sidney Gerst"
date: "November 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read in penguins.csv data

Packages: tidyverse, car, onewaytests

```{r}

library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")

```

Our question is: Is there a signifigant different in mean flipper lengths for female Chinstrap, Gentoo, and Adelie Penguins near Palmer Archipelago?

Maybe I can use a one-way ANOVA...
Assumptions:
- Independent observations
- Equal variance (Levene's Test)
- Normality (histograms/qqplots)

```{r}

counts <- with(penguins, table(species, sex)) #table() great way to look at factors with different values

penguins_2 <- penguins %>%
  mutate(
    common_name = 
      case_when(
        species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
        species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
        species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
      )
  ) %>%
  select(-species) %>%
  filter(sex == "FEMALE")

```

Exploratory graphs and Levene's test

```{r}
hist <- ggplot(penguins_2, aes(x = flipper_length))+
  geom_histogram(aes(fill = common_name))+
  facet_wrap(~ common_name, scale = "free")

hist


qq <- ggplot(penguins_2, aes(sample = flipper_length))+
  geom_qq(aes(color = common_name))+
  facet_wrap(~ common_name, scale = "free")

qq

#based on qq plot and histograms, observations for each group appear normally distributed. also, n >30 for each group, so means will be normally distributed by CLT. 
```

Evaluate variances and performs a Levene's test

```{r}
#if the largest sample variances is <4x greater than the smallest sample variance, then these are "close enough" to consider variances equal

variances <- penguins_2 %>%
  group_by(common_name)%>%
  summarize(
    variance = var(flipper_length)
  )

variances #largest variance (33.1) is deff not more than 4x greater than the smallest variance (15.1)

#Levene's test
#leveneTest function in the 'car' package

levene_penguin <- leveneTest(flipper_length ~ common_name, data = penguins_2)
levene_penguin

#H0: no diff in variances, or variances are equal
#HA: variances are NOT equal

#there is no signifigant difference in variances across species. (p = 0.05782)

```

One-way ANOVA for penguin flipper lengths by species (common_name)

```{r}
penguin_aov <- aov(flipper_length ~ common_name, data = penguins_2)
summary(penguin_aov)

#H0: means across all groups are equal
#HA: at least two group means differ

#post-hoc testing by Tukey's HSD

penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph

#H0: no sig diff in pair-wise means
#HA: sig diff in pair-wise means


#aov.test function in the 'onewaytests' package DOES allow for inline refferences
```
Mean flipper lengths of female penguins differed signigigantly for all species studied (one-way anova, F(df) = statistic, p < 0.001, $\alpha$ = 0.05 with post-hoc Tukey's HSD, $\alpha$ = 0.05).


